### 2015-07-23 LaTeX/skewnormal.tex
updates to plots and manuscript

\section{Spatial skew processes}\label{s:spatialskew}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Genton2004}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

@@

For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
After marginalizing over both $z$ and $\sigma$,
\begin{align}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
\end{align}
that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \text{cor}[v(\bs), v(\bt)] = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.

@@

\section{Data analysis}\label{s:analysis}
To illustrate this method, we consider 31 daily observations of maximum 8-hour ozone measurements for July 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
Initially, we fit a linear regression assuming a mean function of
\begin{align}
  \E[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
\end{align}
The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
\begin{figure}
  \centering
  \includegraphics[width=0.56\linewidth]{plots/ozone-10jul-us.pdf}
  \includegraphics[width=0.41\linewidth]{plots/qq-res.pdf}
  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
  \label{fig:ozone}
\end{figure}
Standard exploratory data analysis techniques for extremal dependence are very challenging with only 31 days worth of data because it is difficult to estimate extreme quantiles at each site to obtain empirical estimates of $\chi$.
Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurences of high sample quantiles.
Figure \ref{fig:bivariateozone} suggests suggests there is more agreement between sites that are close to one another than sites that are far from one another.
\begin{figure}
  \label{fig:bivariateozone}
  \centering
  \includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
  \caption{Daily quantiles for sites that are close (left) and far apart (right)}
\end{figure}

@@

Each chain of the MCMC ran for 30000 iterations with a burn-in period of 25000 iterations.
Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
For each model, Brier scores were averaged over all sites and days to obtain a single Brier score for each dataset.
At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
We then compute the relative (to Gaussian) Brier scores (see Section \ref{s:simresults}) to compare each model.

@@

We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile of the posterior predictive density for July in South Carolina and Georgia.
We fit the model using four methods, two reference and two that performed better.
These four methods are
\begin{enumerate}\setlength{\itemsep}{-0.5em}
  \item Gaussian (reference)
  \item Skew-$t$, $K =$ 1 knot, $T = $ 0, no time series (reference)
  \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series (comparison)
  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series (comparison).
\end{enumerate}
In the bottom two plots, we plot the differences between method 4 and methods 1 and 2.
The most noticeable differences between the reference methods and the comparison methods is that the comparison methods tend to give higher estimates of the 99th quantile along the I-85 corridor between Charlotte and Atlanta.

@@

\hl{NEED TO ADD STUFF HERE ACKNOWLEDGING THAT MARGINALS ARE DIFFERENT}
Also, add an explanation about the purpose of the plot in \ref{fig:bs-marginal}
\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-marginal.pdf}
  \caption{Comparison of Brier score performance by marginal quantile representing 75ppb}
  \label{fig:bs-marginal}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{plots/q99-ozone.pdf}
  \caption{(a) -- (d) give give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, (e) gives the difference between $\widehat{q}(0.99)$ in plots (d) and (a), (f) gives the difference between $\widehat{q}(0.99)$ in plots (d) and (b).}
  \label{fig:ozoneq99}
\end{figure}

@@

\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
Let $N(A)$ be the number of knots in $A$, the area between sites $\bs_1$ and $\bs_2$.
Consider a spatial Poisson process with intensity $\mu(A)$.
So,
\begin{align*}
  P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
\end{align*}
Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.

@@

\subsubsection*{Extremal dependence}
For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
\begin{align} \label{eq:chiskew-t}
  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
\end{align}
where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}(y_1, y_2)$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.

>>>>

\section{Spatial skew processes}\label{s:spatialskew}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Azzalini2014}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

@@

For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
After marginalizing over both $z$ and $\sigma$,
\begin{align}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
\end{align}
that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \text{cor}[v(\bs), v(\bt)] = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.

@@

\section{Data analysis}\label{s:analysis}
To illustrate this method, we consider 31 daily observations of maximum 8-hour ozone measurements for July 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
Initially, we fit a linear regression assuming a mean function of
\begin{align}
  \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
\end{align}
The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
\begin{figure}
  \centering
  \includegraphics[width=0.56\linewidth]{plots/ozone-10jul-us.pdf}
  \includegraphics[width=0.41\linewidth]{plots/qq-res.pdf}
  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
  \label{fig:ozone}
\end{figure}

Standard exploratory data analysis techniques for extremal dependence are very challenging with only 31 days worth of data because it is difficult to estimate extreme quantiles at each site to obtain empirical estimates of $\chi$.
Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurences of high sample quantiles.
Figure \ref{fig:bivariateozone} suggests suggests there is more agreement between sites that are close to one another than sites that are far from one another.
\begin{figure}
  \label{fig:bivariateozone}
  \centering
  \includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
  \caption{Daily quantiles for sites that are close (left) and far apart (right)}
\end{figure}

@@

Each chain of the MCMC ran for 30,000 iterations with a burn-in period of 25,000 iterations.
Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
For each model, Brier scores were averaged over all sites and days to obtain a single Brier score for each dataset.
At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
We then compute the relative (to Gaussian) Brier scores (see Section \ref{s:simresults}) to compare each model.

@@

We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile for South Carolina and Georgia, a subset of the spatial domain, in order to study local features.
The four methods used are
\begin{enumerate}\setlength{\itemsep}{-0.5em}
  \item Gaussian
  \item Skew-$t$, $K =$ 1 knot, $T = $ 0, no time series
  \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series
  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series.
\end{enumerate}
In the bottom two plots, we plot the differences between method 4 and methods 1 and 2.
The most noticeable differences between the reference methods and the comparison methods is that the comparison methods tend to give higher estimates of the 99th quantile along the I-85 corridor between Charlotte and Atlanta.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{plots/q99-ozone.pdf}
  \caption{Panels (a) -- (d) give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, panel (e) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (a), panel (f) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (b).}
  \label{fig:ozoneq99}
\end{figure}

@@

\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
Consider a homogeneous spatial Poisson process with intensity $\mu$.
Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
Let $N(A)$ be the number of points in $A$, and let
\begin{align*}
  \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
\end{align*}
Then
\begin{align*}
  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
\end{align*}
which goes to one as $h \rightarrow \infty$.

% Let $N(A)$ be the number of knots in $A$.
% So,
% \begin{align*}
%   P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
% \end{align*}
% Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
% With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
% Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.

@@

\subsubsection*{Extremal dependence}
For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
\begin{align} \label{eq:chiskew-t}
  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
\end{align}
where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
